Soling at the 2000 Summer Olympics

### 2020-07-27 inst/f1points.Rmd
include zipf law rank comparison

The points scored by each driver give a ranking shown in the table;
but Plackett-Luce likelihoods also give a ranking that does not depend
on any arbitrary points system.  We can plot one ranking against the
other:

```{r, label=calcpoints, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()[[1]]
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

@@

(note carefully that the historically correct points awarded to the
drivers differs from that calculated here.  That is for two reasons:
firstly, ``fastest lap'' points are not included here, and also the
truncation of the order table to the first 11 drivers can increase the
rank of a driver if non-first-11 drivers are placed).

```{r,label=likevspoints,cache=TRUE,echo=FALSE,fig.cap="Ordering of the \\label{orderingbypointsandlikelihood} top-ranked 11 drivers in the 2017 Formula 1 season.  Horizontal axis gives official (points-based) order, and the vertical axis gives the likelihood horder.    Thus, taking 201 as an example,  the points-based ordering would be Hamilton first, then Vettel, then R\"{a}ikk\"{o}nen; YES I KNOW IT IS WRONG: RMARKDOWN IS DEFECTIVE while the likelihood ordering is (reading vertically) Hamilton, Vettel, Bottas ALSO R PLOTTING SUCKS HERE AND I WILL HAVE TO CHANGE THE PDF BY HAND WITH INKSCAPE TO MAKE IT LOOK GOOD"}
par(mfrow=c(2,2))
par(mar = c(3,3,3,3))

@@

From figure \ref{orderingbypointsandlikelihood}, we see a comparison
between the two ordering systems.  Taking 2017 as an example, drivers
Hamilton and Vettel are are respectively first and second according to
both ranking ranking procedures; but R\"{a}ikk\"{o}nen YES I KNOW IT
IS WRONG: RMARKDOWN IS DEFECTIVE and Bottas are third and fourth, and
fourth and third, according to points and likelihood respectively.
The degree of agreement between the two systems might be quantified as
"2", as the positions of the top two drivers agree.  Thus the degree
of agreement for the years 2016-9 would be 8,2,6, and 2 respectively.

>>>>

```{r, label=calcpoints_real, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()[[1]]   # real points system
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

@@

```{r,label=likevspoints,cache=TRUE,echo=FALSE,fig.cap="Ordering of the \\label{orderingbypointsandlikelihood} top-ranked 11 drivers in Formula 1 seasons 2016-2019.  Horizontal axis gives official (points-based) order, and the vertical axis gives the likelihood horder.    Thus, taking 2017 as an example,  the points-based ordering would be Hamilton first, then Vettel, then R\"{a}ikk\"{o}nen; YES I KNOW IT IS WRONG: RMARKDOWN IS DEFECTIVE while the likelihood ordering is (reading vertically) Hamilton, Vettel, Bottas ALSO R PLOTTING SUCKS HERE AND I WILL HAVE TO CHANGE THE PDF BY HAND WITH INKSCAPE TO MAKE IT LOOK GOOD"}
par(mfrow=c(2,2))
par(mar = c(3,3,3,3))

@@

The points scored by each driver give a ranking shown in the table;
but Plackett-Luce likelihoods also give a ranking that does not depend
on any arbitrary points system.  We can plot one ranking against the
other, see Figure \ref{orderingbypointsandlikelihood}\footnote{The
historically correct points awarded to the drivers differs from that
calculated here.  That is for two reasons: firstly, ``fastest lap''
points are not included here, and also the truncation of the order
table to the first 11 drivers can increase the rank of a driver if
non-first-11 drivers are placed}.  We thus see a comparison between
the two ordering systems.  Taking 2017 as an example, drivers Hamilton
and Vettel are are respectively first and second according to both
ranking ranking procedures; but R\"{a}ikk\"{o}nen YES I KNOW IT IS
WRONG: RMARKDOWN IS DEFECTIVE and Bottas are third and fourth, and
fourth and third, according to points and likelihood respectively.
The degree of agreement between the two systems might be quantified as
"2", as the positions of the top two drivers agree.  Thus the degree
of agreement between likelihood ranking and points ranking for the
years 2016-9 would be 8,2,6, and 2 respectively.

However, the points system used is essentially arbitrary.  We could
use, for example, Zipf's law to rank the drivers: award one point to
the winner, half a point to second place, one third of a point to
third, and so on; see figure \ref{orderingbypointsandlikelihood_zipf}
in which we see generally poorer agreement between points-based ranks
and likelihood-based ranks.

```{r, label=calcpoints_zipf, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()$zipf
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

```{r,label=likevspointszipf,cache=TRUE,echo=FALSE,fig.cap="As for figure \ref{likevspoints} \\label{orderingbypointsandlikelihood_zipf} but points ranking calculated by Zipf's law.  Note the generally poorer agreement between the two ranking systems ALSO R PLOTTING SUCKS HERE AND I WILL HAVE TO CHANGE THE PDF BY HAND WITH INKSCAPE TO MAKE IT LOOK GOOD"}
par(mfrow=c(2,2))
par(mar = c(3,3,3,3))
